### 2018-05-25 abstract.tex
Fix syntax errors according to the first revision by Prof. Xue

\begin{abstract}
  GEneral Matrix Multiply (GEMM) is the most fundamental
  computational kernel routine in BLAS library.
  To achieve best performance, in-memory data must be
  prefetched into fast on-chip caches ahead of use.
  Two techniques, software prefetching and data packing,
  are utilized to effectively exploit the capability of on-chip caches.
  These techniques are sufficient for lastly-recent-used (LRU) caches,
  which is often the case in traditional high-performance
  processors used in high-end servers and supercomputers.
  But in recent years the market is meeting a diversity
  in processor design, and some high-performance processors
  are equipped with shared caches with non-LRU replacement policies.
  This rises a challenge to GEMM development in multi-threaded context.
  As several threads try to load data into the shared cache simultaneously,
  inter-thread cache conflicts would increase significantly.
  We present a Shared Cache Partition (SCP) method to
  eliminate inter-thread cache conflicts in GEMM routines,
  by partitioning the shared cache into physically disjoint
  sets and assign different sets to different threads.
  We implement SCP in the OpenBLAS library and
  evaluate the performance on Phytium 2000+,
  a 64-core AArch64 processor with private LRU L1 caches
  and shared (by every 4 cores) psudo-random L2 caches.
  The results show that conflict misses on both L1 and L2
  caches are effectively reduced and GEMM performance get improved
  by $2.75\%$--$6.91\%$.
  %% FIXME BLIS? FT1500A?
\end{abstract}

>>>>

\begin{abstract}
  GEneral Matrix Multiply (GEMM) is the most fundamental
  computational kernel routine in the BLAS library.
  To achieve high performance, in-memory data must be
  prefetched into fast on-chip caches before they are used.
  Two techniques, software prefetching and data packing,
  are utilized to effectively exploit the capability of on-chip caches.
  These techniques are sufficient for lastly-recent-used (LRU) caches,
  which are often used in high-performance
  processors deployed in high-end servers and supercomputers.
  But in recent years the market has become increasingly more diverse
  in processor design, resulting in high-performance processors
  equipped with shared caches with non-LRU replacement policies.
  This poses a challenge to GEMM development in a multi-threaded context.
  As several threads try to load data into the shared cache simultaneously,
  inter-thread cache conflicts will increase significantly.
  We present a Shared Cache Partitioning (SCP) method to
  eliminate inter-thread cache conflicts in GEMM routines,
  by partitioning the shared cache into physically disjoint
  sets and assign different sets to different threads.
  We implement SCP in the OpenBLAS library and
  evaluate the performance on Phytium 2000+,
  a 64-core AArch64 processor with private LRU L1 caches
  and shared (by every 4 cores) psudo-random L2 caches.
  The results show that conflict misses on both L1 and L2
  caches are effectively reduced, resulting in an improvement
  of the GEMM performance by $2.75\%$--$6.91\%$.
\end{abstract}

### 2018-05-25 background.tex
Fix syntax errors according to the first revision by Prof. Xue

Modern architectures take advantage of fast on-chip caches
to overcome the ``memory wall''.
As a compromise between direct mapped cache and full-associative cache,
set-associative cache has been the dominant choice for real-world architectures.
To understand how shared cache can cause inter-thread cache conflicts
and do harm to GEMM performance, we review some details
about the working mechanisim of set-associative caches.

A set-associative cache can be defined by a 4-tuple $(c, l, n, nt)$,
in which $c$ is the cache capacity, $l$ is the cache line size,
$n$ is ways-of-associativity,
and $nt$ means the number of cores sharing the cache.
The whole cache memory is divided into $ns=c/l/n$ sets,
denoted by integral indices $\mathcal{S} = [0, ns)$,
with each set containing $n$ cache lines.
Generally, $ns$ is a power-of-two value.
An indexing function $\varphi: \mathcal{A} \mapsto \mathcal{S}$ is responsible for
mapping addresses from the address space $\mathcal{A}$
(either virtual or physical) to cache sets $\mathcal{S}$.
Data at $addr \in \mathcal{A}$ must be fetched
into the cache set $\varphi(addr) \in \mathcal{S}$ if requested.
A classic implementation of the indexing function is shown below,
%% $\varphi(addr) = (addr \gg log_2(l)) ~\&~ (ns-1)$,
with $\gg$ and $\&$ representing logical-shift-right and bitwise-and operations, respectively.
It is an extraction of the $log_2(ns)$ bits starting from the $log_2(l)$
position in $addr$'s binary form.
\begin{equation*}
  \varphi(addr) = (addr \gg log_2(l)) ~\&~ (ns-1)
  \label{eq:phi}
\end{equation*}

@@

Throughout this paper we will use a 4-core processor with
$NL=2$ levels of caches as the platform for SCP demonstration.
Figure.~\ref{fig:hierachy} shows its memory hierachy and
detailed architectural parameters are listed in Table.~\ref{tab:cluster}.
Each core has 32 128-bit (16B) vector registers,
each capable of store 2 double precision floating-point numbers.
To make a consistency, the processor cores on level 0
is represented by its registers, which can be viewed as a special
level-0 cache $L0 = (512B, 16B, 32, 1)$ with programmable replacement policy.

@@

In this blocked algorithm, the loops over the $N$, $K$ and $M$
matrix dimensions are tiled by sizes $N_c$, $K_c$ and $M_c$ at 
layers 1, 2 and 3, respectively.
At layers 4 and 5, the $N$ and $M$ dimensions are further tiled by sizes
$N_r$ and $M_r$, respectively.
As a result, the innermost loop at layer 6 goes over the $K$
dimension for a total of $K_c$ times,
with each iteration performing a rank-1 update on
the $M_r \times N_r$ submatrix of $C$, as shown at layer 7.
In OpenBLAS, matrix $C$ is scaled by $\beta$ before the loops,
so the multiply-accumulate operation at layer 7 need not to scale $C$ again.
Tiling factors $N_c$, $K_c$, $M_c$, $N_r$ and $M_r$ are carefully selected so that
matrices on each level fit into a certain level in the memory hierachy.
Equation~(\ref{eq:constraints.reg})--(\ref{eq:constraints.l3}) show
the constraints on these tiling factors.
$es$ means the size of matrix element, e.g. 8B for double precision floating-point numbers,
and $c_l$ means the capacity of registers or caches on level $l$.
First, by (\ref{eq:constraints.reg}), $M_r$ and $N_r$ are constrained that $M_r$ elements from $A$,
$N_r$ elements from $B$ and $M_r \times N_r$ elements from $C$ can fit into registers
(the psudo level-0 cache).
Then, by (\ref{eq:constraints.l1}), $K_c$ is constrained that $B_4$ ($N_r \times K_c$),
$A_4$ ($M_r \times K_c$), and $A_4$'s counterpart in next iteration
at layer 5 fit into the L1 cache.
Then, by (\ref{eq:constraints.l2}), $M_c$ is constrained that $A_3$ ($M_c \times K_c$),
$B_3$ ($N_r \times K_c$), and $B_3$'s counterpart in next iteration
at layer 4 fit into the L2 cache.
At last, by (\ref{eq:constraints.l3}), $N_c$ is constrained that 
$B_2$ ($K_c \times N_c$) and $A_2$ ($M_c \times K_c$) fit into the L3 cache, if it exists.
Larger matrices on higher layers 1 -- 3 live in main memory.

@@

As stated in Section~\ref{sec:intro}, an GEMM implementation
is made up by a kernel routine and an overall strategy.
In Figure.~\ref{fig:gemm}, GEBP surves as the kernel routine,
and the overall strategy is defined by the three outter loops at layers 1 -- 3.
Developer can control how to partition the workload by choosing
tile factors $M_c$, $N_c$ and $K_c$,
and how to schedule GEBP tasks by interchanging loop orders at layers 1 - 3.
Either the N loop at layer 1 or M loop at layer 3, or both of them,
can be parallelized in a multi-threaded context.
Specifically, the strategy in Figure.~\ref{fig:gemm} chooses a NKM loop order and
parallelizes the M loop at layer 3.
Different threads work on different parts of $A_1$
and share the same $B_1$.
As a result, there are multiple $A_2$ instances, one per thread,
and a single $B_2$ instances, shared by all threads.
To understand how the loop on layer 3 is parallelized on the 4-core processor, 
Figure.~\ref{fig:workload} shows the data accessing patterns of a single thread $T_1$.
While all color shaded submatrices are accessed by $T_1$,
only z-curve masked submatrices (a subset of color shaded ones)
are packed by the thread.
That is, each thread packs its own $A_2$ instance and 
$B_2$ is packed by all thread collabratively.
%% We will use the OpenBLAS GEMM implementation in Figure.~\ref{fig:gemm}
%% in discussion and evaluation throughout the paper.

@@

For set-associative caches, cache misses can be categorized
into capacity misses and conflict misses.
While constraints (\ref{eq:constraints.reg}) -- (\ref{eq:constraints.l3})
serve to avoid capacity misses,
conflict misses can not be eliminated completely.
%% We use the term ``cache conflict'' to represent conflict miss events.

Recall Figure.~\ref{fig:gemm}, the packed matrix $A_2$
will be read $N_c/N_r$ times within a single GEBP execution,
and is expected to reside in L2 cache during all $N_c/N_r$ iterations at layer 4.
The packed matrix $B_4$ is similar that it is expected to
reside in L1 cache during all $M_c/M_r$ iterations at layer 5.
In general, even with carefully selected tile factors,
matrix data can still be evicted out for a few reasons.
For instance, data structures other than the packed matrices,
or program code (in case of a unified cache) may be fetched into a
cache set the matrix data lives in.
We call this kind of cache conflicts intra-thread cache conflicts
because it occurs within the execution of a single thread.
To deal with intra-thread conflicts,
the GEBP kernel repeatedly prefetch data from $A_4$ and $B_4$ to L1 cache
in every iteration at layer 6 before it actually reads them in case of
the data being evicted out accidentally.
As GEBP accesses the matrix data much more frequently than other
data structures, this repeated prefething stratecy
can remedy the penalty caused by intra-thread conflicts.

@@

In GEMM, the cache is utilized aggressively
that almost all cache space is expected to be occupied by packed matrices.
Prefetch instructions from different threads are executed repeatedly,
interleaved in an unpredictable manner.
So inter-thread cache conflicts described above
can occur frequently on non-LRU shared caches,
thus making the technique dealing with intra-thread cache conflicts helpless.
The situation goes worse with more threads sharing the same cache.

>>>>

Modern architectures take advantage of fast on-chip caches
to overcome the ``memory wall''.
As a compromise between direct mapped cache and full-associative cache,
set-associative cache has been the dominant choice for real-world architectures.
To understand how shared cache can cause inter-thread cache conflicts
and hurts GEMM performance, we review some details
about the working mechanisim of set-associative caches.

A set-associative cache can be defined by a 4-tuple $(c, l, n, nt)$,
in which $c$ is the cache capacity, $l$ is the cache line size,
$n$ is ways-of-associativity,
and $nt$ means the number of cores sharing the cache.
The whole cache memory is divided into $ns=c/l/n$ sets,
denoted by integral indices $\mathcal{S} = [0, ns)$,
with each set containing $n$ cache lines.
Generally, $ns$ is a power-of-two.
An indexing function $\varphi: \mathcal{A} \mapsto \mathcal{S}$ is responsible for
mapping addresses from the address space $\mathcal{A}$
(either virtual or physical) to cache sets $\mathcal{S}$.
Data at $addr \in \mathcal{A}$ must be fetched
into the cache set $\varphi(addr) \in \mathcal{S}$ if requested.
A classic implementation of the indexing function is shown below,
%% $\varphi(addr) = (addr \gg log_2(l)) ~\&~ (ns-1)$,
with $\gg$ and $\&$ representing logical-shift-right and bitwise-and operations, respectively.
It is an extraction of the $log_2(ns)$ bits starting from the $log_2(l)$
position in $addr$'s binary form.
\begin{equation*}
  \varphi(addr) = (addr \gg log_2(l)) ~\&~ (ns-1)
  \label{eq:phi}
\end{equation*}

@@

Throughout this paper we will use a 4-core processor with
$NL=2$ levels of caches as the platform for SCP demonstration.
Figure.~\ref{fig:hierachy} shows its memory hierachy and
detailed architectural parameters are listed in Table.~\ref{tab:cluster}.
Each core has 32 128-bit (16B) vector registers,
each capable of storing 2 double precision floating-point numbers.
For convenience, the processor cores on level 0
is represented by its registers, which can be viewed as a special
level-0 cache $L0 = (512B, 16B, 32, 1)$ with programmable replacement policy.

@@

In this blocked algorithm, the loops over the $N$, $K$ and $M$
matrix dimensions are tiled by sizes $N_c$, $K_c$ and $M_c$ at 
layers 1, 2 and 3, respectively.
At layers 4 and 5, the $N$ and $M$ dimensions are further tiled by sizes
$N_r$ and $M_r$, respectively.
As a result, the innermost loop at layer 6 goes over the $K$
dimension for a total of $K_c$ times,
with each iteration performing a rank-1 update on
the $M_r \times N_r$ submatrix of $C$, as shown at layer 7.
In OpenBLAS, matrix $C$ is scaled by $\beta$ before the loops,
so the multiply-accumulate operation at layer 7 does not have to scale $C$ again.
Tiling factors $N_c$, $K_c$, $M_c$, $N_r$ and $M_r$ are carefully selected so that
matrices on each level fit into a certain level in the memory hierachy.
Equation~(\ref{eq:constraints.reg})--(\ref{eq:constraints.l3}) show
the constraints on these tiling factors.
$es$ denotes the size of matrix element, e.g. 8B for double precision floating-point numbers,
and $c_l$ denotes the capacity of registers or caches on level $l$.
First, by (\ref{eq:constraints.reg}), $M_r$ and $N_r$ are constrained that $M_r$ elements from $A$,
$N_r$ elements from $B$ and $M_r \times N_r$ elements from $C$ can fit into registers
(the psudo level-0 cache).
Then, by (\ref{eq:constraints.l1}), $K_c$ is constrained that $B_4$ ($N_r \times K_c$),
$A_4$ ($M_r \times K_c$), and $A_4$'s counterpart in the next iteration
at layer 5 fit into the L1 cache.
Then, by (\ref{eq:constraints.l2}), $M_c$ is constrained that $A_3$ ($M_c \times K_c$),
$B_3$ ($N_r \times K_c$), and $B_3$'s counterpart in the next iteration
at layer 4 fit into the L2 cache.
At last, by (\ref{eq:constraints.l3}), $N_c$ is constrained that 
$B_2$ ($K_c \times N_c$) and $A_2$ ($M_c \times K_c$) fit into the L3 cache, if it exists.
Larger matrices on higher layers 1 -- 3 always reside in main memory.

@@

As stated in Section~\ref{sec:intro}, an GEMM implementation
is made up by a kernel routine and an overall strategy.
In Figure.~\ref{fig:gemm}, GEBP surves as the kernel routine,
and the overall strategy is defined by the three outter loops at layers 1 -- 3.
Developers can control how to partition the workload by choosing
tile factors $M_c$, $N_c$ and $K_c$,
and how to schedule GEBP tasks by interchanging loop orders at layers 1 - 3.
Either the N loop at layer 1 or M loop at layer 3, or both of them,
can be parallelized in a multi-threaded context.
Specifically, the strategy in Figure.~\ref{fig:gemm} chooses a NKM loop order and
parallelizes the M loop at layer 3.
Different threads work on different parts of $A_1$
and share the same $B_1$.
As a result, there are multiple $A_2$ instances, one per thread,
and a single $B_2$ instances, shared by all threads.
To understand how the loop on layer 3 is parallelized on the 4-core processor, 
Figure.~\ref{fig:workload} shows the data accessing patterns of a single thread $T_1$.
While all color shaded submatrices are accessed by $T_1$,
only z-curve masked submatrices (a subset of color shaded ones)
are packed by the thread.
That is, each thread packs its own $A_2$ instance and 
$B_2$ is packed by all thread collabratively.
%% We will use the OpenBLAS GEMM implementation in Figure.~\ref{fig:gemm}
%% in discussion and evaluation throughout the paper.

@@

For set-associative caches, cache misses can be categorized
into capacity misses and conflict misses.
While constraints (\ref{eq:constraints.reg}) -- (\ref{eq:constraints.l3})
serve to avoid capacity misses,
conflict misses cannot be eliminated completely.
%% We use the term ``cache conflict'' to represent conflict miss events.

In Figure.~\ref{fig:gemm}, the packed matrix $A_2$
will be read $N_c/N_r$ times within a single GEBP execution,
and is expected to reside in L2 cache during all $N_c/N_r$ iterations at layer 4.
The packed matrix $B_4$ is similar in that it is expected to
reside in L1 cache during all $M_c/M_r$ iterations at layer 5.
In general, even with carefully selected tile factors,
matrix data can still be evicted out for a few reasons.
For instance, data structures other than the packed matrices,
or program code (in case of a unified cache) may be fetched into a
cache set the matrix data lives in.
We call this kind of cache conflicts intra-thread cache conflicts
because it occurs within the execution of a single thread.
To deal with intra-thread conflicts,
the GEBP kernel repeatedly prefetch data from $A_4$ and $B_4$ to L1 cache
in every iteration at layer 6 before it actually reads them in case of
the data being evicted out accidentally.
As GEBP accesses the matrix data much more frequently than other
data structures, this repeated prefething strategy
can reduce the penalty caused by intra-thread conflicts.

@@

In GEMM, the cache is utilized aggressively
that almost all cache space is expected to be occupied by packed matrices.
Prefetch instructions from different threads are executed repeatedly,
interleaved in an unpredictable manner.
So inter-thread cache conflicts described above
can occur frequently on non-LRU shared caches,
thus making the technique dealing with intra-thread cache conflicts ineffective.
The situation worsens with more threads sharing the same cache.

### 2018-05-25 evaluation.tex
Fix syntax errors according to the first revision by Prof. Xue

We implement SCP in the OpenBLAS~\cite{openblas} library version 0.3.0-dev,
and evaluate it on a Phytium 2000+ processor.
The Phytium 2000+ processor is an emerging high-performance
64-core processor based on ARM's AArch64 architecture.
The 64 cores are organized into 16 clusters with each
cluster containing 4 cores.
The structure of the cluster is almost the same with
the contrived 4-core processor in Figure.~\ref{fig:hierachy}
except that all 16 clusters connect to the main memory
with their L2 caches coherented by hardare.
See Table.~\ref{tab:cluster} for architectural parameters of the cores and caches.
We restrict our evaluation to DGEMM,
as in prior work~\cite{blispar,augem,poetmicro}, for two reasons.
First, the basic idea behind SCP applies to other
variants of GEMM such as SGEMM, CGEMM and ZGEMM.
Second, the LINPACK benchmark, which is used to build the
TOP500 list of world's most powerful supercomputers (FIXME citation),
relys on the DGEMM variant.

All evaluations run with the same tiling factors
$M_r$, $N_r$, $K_c$, $M_c$ and $N_c$.
The tiling factors are computed as in Section~\ref{subsec:example},
i.e. $M_r = 4$, $N_r = 8$, $K_c = 256$, $M_c = 192$, $N_c = 1024NT$.
The parallelism is controlled by two parameters,
the number of active clusters $NC$,
and the number of active threads per cluster $NT_C$.
The total number of threads is $NT = NC \cdot NT_C$.
For the Phytium 2000+ processor, the valid ranges for $NC$ and $NT_C$
are $NC \in [1, 16]$ and $NT_C \in [1, 4]$, respectively.
Generally, DGEMM performance are presented in $flops$.
%% floating-point operations per second ($flops$).
As $NT$ varies in our evaluation,
we use another metric derived from $flops$,
the average thread efficiency,
to present performance results consistently.
The average thread efficiency is a normalized value computed as
$E_{avg} = flops / (NT \cdot \widehat{flops})$
in which $\widehat{flops}$ stands for the theoretical performance peak of a single core.
%% \begin{equation*}
%%   E = flops / (NT \cdot \widehat{flops})
%% \end{equation*}

We first show how cache sharing can harm GEMM performance
in Section~\ref{subsec:drawback}.
Then Section~\ref{subsec:benefit} demonstrates how SCP can remedy
the penalty caused by cache sharing.
Quantitative analysis of the cache miss rates
is presented in Section~\ref{subsec:analysis}.
Finally, Section~\ref{subsec:privb} discusses
the privatization of shared matrix $B_2$.
%% Three versions of DGEMM are evaluated:
%% (1) Base, the baseline version,
%% (2) SCP,  the SCP version, 
%% and (3) SCP-P, the SCP version plus privatization of shared matrix $B_2$.

@@

\begin{table}
  \centering
  \caption{Average thread efficiency win of SCP}
  \label{tab:win}
  \setlength{\tabcolsep}{3.5pt}
  \begin{tabular}{cccccc}
    \toprule
     & $NT=4$ & $NT=8$ & $NT=16$ & $NT=32$ & $NT=64$ \\
    \midrule
    $NT_C=2$ & 1.77 & 1.69 & 1.82 & 1.78 & - \\
    $NT_C=4$ & 2.75 & 4.62 & 4.41 & 3.11 & 6.91 \\
    \bottomrule
  \end{tabular}
\end{table}

@@

Under all parallelism configurations, 
SCP performs better than the Base
consistently over all matrix sizes.
Table.~\ref{tab:win} summarizes the performance win of SCP over Base.
%% Table.~\ref{tab:benefit} summarizes the average value
%% of $E_{avg}$ in SCP evaluation.
%% There are two values in each cell, separated by a slash.
%% $E_{avg}$ stands on the left,
%% and performance win over conventional packing
%% in the corresponding parallelism configuration
%% is shown on the right for convenience.
With $NT_C=2$, SCP performs slightly better than Base
by $1.7\%$--$1.8\%$.
The performance gap becomes larger with $NT_C=4$.
ranging from $2.75\%$ to $6.91\%$.
The largest performance gain, $6.91\%$, which is a considerable improvement,
is observed under the maximum parallelism $NT=64$.
%% Looking at Table.~\ref{tab:benefit} vertically,
%% we can see that SCP still cannot fully eliminate the impact
%% of cache sharing.
%% But it do relieve the situation to a certain degree.

@@

\subsection{Privatization of Shared Matrix}\label{subsec:privb}
In previous evaluations, SCP only set-partitions the shared cache
for $A_2$ matrices which are private to threads.
The shared $B_2$ matrix is still packed
in conventional way-partition style.
As mentioned in Section~\ref{subsec:example},
set-partition can also be applied to $B_2$ after $B_2$ is privatized.
In this section, we will discuss and evaluate this privatization alternative.

>>>>

We implement SCP in the OpenBLAS~\cite{openblas} library version 0.3.0-dev,
and evaluate it on a Phytium 2000+ processor.
The Phytium 2000+ processor is an emerging high-performance
64-core processor based on ARM's AArch64 architecture.
The 64 cores are organized into 16 clusters with each
cluster containing 4 cores.
The structure of the cluster is almost the same with
the contrived 4-core processor in Figure.~\ref{fig:hierachy}
except that L2 caches of all 16 clusters connect to the main memory
with hardware coherence.
Table.~\ref{tab:cluster} lists architectural parameters of the cores and caches.
We restrict our evaluation to DGEMM,
as in prior work~\cite{blispar,augem,poetmicro}, for two reasons.
First, the basic idea behind SCP applies to other
variants of GEMM such as SGEMM, CGEMM and ZGEMM.
Second, the LINPACK benchmark, which is used to build the
TOP500 list of world's most powerful supercomputers (FIXME citation),
relies on the DGEMM variant.

All evaluations run with the same tiling factors
$M_r$, $N_r$, $K_c$, $M_c$ and $N_c$.
The tiling factors are computed as in Section~\ref{subsec:example},
i.e. $M_r = 4$, $N_r = 8$, $K_c = 256$, $M_c = 192$, $N_c = 1024NT$.
The parallelism is controlled by two parameters,
the number of active clusters $NC$,
and the number of active threads per cluster $NT_C$.
The total number of threads is $NT = NC \cdot NT_C$.
For the Phytium 2000+ processor, the valid ranges for $NC$ and $NT_C$
are $NC \in [1, 16]$ and $NT_C \in [1, 4]$, respectively.
Generally, DGEMM performance is presented in $flops$.
%% floating-point operations per second ($flops$).
As $NT$ varies in our evaluation,
we use another metric derived from $flops$,
the average thread efficiency,
to present performance results consistently.
The average thread efficiency is a normalized value computed as
$E_{avg} = flops / (NT \cdot \widehat{flops})$
in which $\widehat{flops}$ stands for the theoretical performance peak of a single core.
%% \begin{equation*}
%%   E = flops / (NT \cdot \widehat{flops})
%% \end{equation*}

We first show how cache sharing can hurt GEMM performance
in Section~\ref{subsec:drawback}.
Then Section~\ref{subsec:benefit} demonstrates how SCP can reduce
the penalty caused by cache sharing.
Quantitative analysis of the cache miss rates
is presented in Section~\ref{subsec:analysis}.
Finally, Section~\ref{subsec:privb} discusses
the privatization of shared matrix $B_2$.
%% Three versions of DGEMM are evaluated:
%% (1) Base, the baseline version,
%% (2) SCP,  the SCP version, 
%% and (3) SCP-P, the SCP version plus privatization of shared matrix $B_2$.

@@

\begin{table}
  \centering
  \caption{Average thread efficiency improvement of SCP}
  \label{tab:win}
  \setlength{\tabcolsep}{3.5pt}
  \begin{tabular}{cccccc}
    \toprule
     & $NT=4$ & $NT=8$ & $NT=16$ & $NT=32$ & $NT=64$ \\
    \midrule
    $NT_C=2$ & 1.77 & 1.69 & 1.82 & 1.78 & - \\
    $NT_C=4$ & 2.75 & 4.62 & 4.41 & 3.11 & 6.91 \\
    \bottomrule
  \end{tabular}
\end{table}

@@

Under all parallelism configurations, 
SCP performs better than the Base
consistently over all matrix sizes.
Table.~\ref{tab:win} summarizes the performance improvement of SCP over Base.
%% Table.~\ref{tab:benefit} summarizes the average value
%% of $E_{avg}$ in SCP evaluation.
%% There are two values in each cell, separated by a slash.
%% $E_{avg}$ stands on the left,
%% and performance win over conventional packing
%% in the corresponding parallelism configuration
%% is shown on the right for convenience.
With $NT_C=2$, SCP performs slightly better than Base
by $1.7\%$--$1.8\%$.
The performance gap becomes larger with $NT_C=4$.
ranging from $2.75\%$ to $6.91\%$.
The largest performance gain, $6.91\%$, which is a considerable improvement,
is observed under the maximum parallelism $NT=64$.
%% Looking at Table.~\ref{tab:benefit} vertically,
%% we can see that SCP still cannot fully eliminate the impact
%% of cache sharing.
%% But it do relieve the situation to a certain degree.

@@

\subsection{Privatization of Shared Matrix}\label{subsec:privb}
In previous evaluations, SCP only set-partitions the shared cache
for $A_2$ matrices which are private to threads.
The shared $B_2$ matrix is still packed
in conventional way-partitioning style.
As mentioned in Section~\ref{subsec:example},
set-partitioning can also be applied to $B_2$ after $B_2$ is privatized.
In this section, we will discuss and evaluate this privatization alternative.
